Drake Plan

### 2019-08-22 AOD_CONUS_Report.Rmd
CONUS model using one-year data done

---
title: "AOD CONUS Modeling Results"
author: "Yang Liu"
date: "08/17/2019"
output: html_document
---

```{r}
library(drake) 
loadd(plotAOD, cache = ssd_cache)
n_vars = length(plotAOD)
fig_height = ceiling(n_vars/2)*3.5
```

@@

```{r, fig.width = 10, fig.height = fig_height}
loadd(plotAOD, cache = ssd_cache)
gridExtra::grid.arrange(grobs = plotAOD, ncol = 2)
```

@@

```{r}
# top features
loadd("cv_results_new_var_join_mcd_terra", cache = ssd_cache)
# rferesults = cv_results_new_var_join_mcd_terra
cv_results_new_var_join_mcd_terra$features_rank_rfe
```

@@

```{r}
# top features
loadd("cv_results_new_var_join_mcd_aqua", cache = ssd_cache)
cv_results_new_var_join_mcd_aqua$features_rank_rfe
```

>>>>

---
title: "AOD CONUS Modeling Results"
author: "Yang Liu"
date: "08/17/2019"
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
```

```{r}
library(drake) 
loadd(plotscatter, cache = ssd_cache)
n_vars = length(plotscatter)
fig_height = ceiling(n_vars/2)*3.5
```

@@

```{r, fig.width = 10, fig.height = 7}
loadd(plotAOD, cache = ssd_cache)
gridExtra::grid.arrange(grobs = plotAOD, ncol = 2)
```

@@

```{r}
# top features
loadd("cv_results_terra", cache = ssd_cache)
# rferesults = cv_results_new_var_join_mcd_terra
cv_results_terra$features_rank_rfe
```

@@

```{r}
# top features
loadd("cv_results_aqua", cache = ssd_cache)
cv_results_aqua$features_rank_rfe
```

### 2019-08-22 R/drake_plan_conus.Rmd
CONUS model using one-year data done

```{r}
# See `drake_funcs.R` for functions used in drake plan #

@@

# Libraries
suppressPackageStartupMessages({
  library(sf)
  library(magrittr)
  library(drake)
  library(data.table)
  library(fst)
  library(future)
  library(here)
  library(nngeo)
  library(ggplot2)
  library(SHAPforxgboost)
})
# ** need to revise accordingly the path to your local Just_universal
source("~/liuyanguu/Just_universal/code/xgboost_cvtune.R")
source("~/liuyanguu/Just_universal/code/xgboost_cv_RFE.R")

# Parameters ####

aer_stn_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/aeronet_locations_v3.txt"

aer_data_path = '~/qnap_geo/aeronet/All_Sites_Times_All_Points_AOD20_20180603.dat'
aer_data_path = '/scratch/temp/All_Sites_Times_All_Points_AOD20_20180603.dat' # don't have time to play with Ethernet all day
# the aer data with AOD470nm
aer_data_path_new = "~/Intermediate/AOD_data/190628_aod20_wPred_NEMIA_idLST.fst"

aer_files_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
date_start = "2018-01-01" # do not include in sel_data_bytime if don't want to limit
date_end   = "2018-12-31" # do not include in sel_data_bytime if don't want to limit
geo_region = "NEMIA"
sats = c("terra", "aqua")
# MCD19 data: 
mcd19path = "/data-coco/mcd19/fst/nemia/2018/" # now there is only 2018 data 
mcd19path_CONUS = "/data-coco/mcd19/fst/conus/2018/"
# point to cache: 
ssd_cache = new_cache(path = "/scratch/cache/aeronet_drake")

# for reporting purpose
y_var = "diff_AOD"
y_var_pred = "diff_AOD_pred"
# how many features to print in scatterplot
n_vars = 4

# CONUS Drake Plans  ------------------------------------------------------------
source(here("R/drake_funcs.R"))
nemia_plan <- drake_plan(
  # 1.  Aeroent ---------------------------------------------------------------------

@@

  # near MODIS cells within 300km
  refsub = target(get_near_cellsid(conus_aer, sel_aer_region, refgrid),
                  transform = map(sel_aer_region, .id = FALSE)),

@@

  # 3. WPred Join MCD19 -----------------------------------------------------------
  # mcd is the large MCD19 dataset, limit to `refsub`:nearby(300km) when readin to 
  # control total lines, unable to read in all data together, too many rows
  mcd = target(read_mcd19(sat, file_in(mcd19path_CONUS)), transform = map(sat = c("terra", "aqua"))),
  # join is the rolling-joined dataset with aeronet
  join = target(rolling_join(mcd_sat = mcd, aer_data_wPred = wPred),
                transform = cross(mcd, wPred, .id = mcd)),

  # select aer sites in mcd19. further limit the stations
  sel_aer = target(aer_in_mcd19(aer_nearest_2, mcd19 = join),
             transform = map(join)),
  # buffers 
  ref = target(ref_in_buffer(sel_aer, refgrid),
             transform = map(sel_aer)),
  # sites
  p = target(get_site_list(region_aer_m = sel_aer), transform = map(sel_aer)),
  n = target(get_site_name(region_aer_m = sel_aer), transform = map(sel_aer)),

  # mapping 
  # this way doesn't work:
  # sub_grid = target(select_refgrid_subset(ref_list0 = ref, point0 = p, name0 = n),
  #              transform = map(ref, p, n))
  sub_grid = target(purrr_pmap(ref = ref, p = p, n = n),
               transform = map(ref, p, n, .id = FALSE)),
  # calculate new vars to obtain the dataset ready for modeling for terra and aqua, using join, mcd, sub_grid 
  new_var = target(aod_MODIS_newVars(aod_join_MODIS = join, 
                                 MODIS_all = mcd, refDT_sub = sub_grid),
               transform = map(join, mcd, sub_grid, .id = c(join))),

  # 5. CV -----------------------------------------------------------
  cv_results = target(run_cv(dt = new_var),transform = map(new_var)),

@@

  shap_long = target(get_SHAP_long(cv_results),
                  transform = map(cv_results,.id = FALSE)),
  plotSHAP1 = target(plot_SHAP(shap_long),
                 transform = map(shap_long,.id = FALSE)),
  plotSHAP2 = target(plot_SHAP_scatter(cv_results, shap_long),
                 transform = map(cv_results, shap_long,.id = FALSE)),
  report = rmarkdown::render(
    knitr_in("AOD_CONUS_Report.Rmd"),
    output_file = file_out("AOD_CONUS_Report.html"))
)
nemia_plan
#drake_plan_source(nemia_plan)
nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
vis_drake_graph(nemia_config, ncol_legend = 0)

make(nemia_plan, cache = ssd_cache)
```

# Run the plan

```{r, eval = F}
make(nemia_plan, cache = ssd_cache)

future::plan(future::multiprocess)
options(future.globals.maxSize= 100*1024^3) 
make(nemia_plan, cache = ssd_cache,
     parallelism = "future", jobs = 20, cache_log_file = TRUE)

## Some simple analysis 
loadd("new_var_join_mcd_terra", cache = ssd_cache)
loadd("new_var_join_mcd_aqua", cache = ssd_cache)
SHAPforxgboost::scatter.plot.diagonal(new_var_join_mcd_aqua, "Optical_Depth_047",
                                      "Mean_AOD270km")
cor(as.matrix(new_var_join_mcd_terra[,c(grep("Mean_AOD", names(new_var_join_mcd_terra), value = T), "AOD_470nm"), with = F]))
## 
loadd(cache = ssd_cache) # load all object

# to see the updated graph, run config again
nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
vis_drake_graph(nemia_config)

>>>>

```{r}
# See `drake_funcs.R` for functions used in this drake plan #

@@

# Libraries
suppressPackageStartupMessages({
  library(sf)
  library(magrittr)
  library(drake)
  library(data.table)
  library(fst)
  library(future)
  library(here)
  library(nngeo)
  library(ggplot2)
  library(SHAPforxgboost)
})
# ** need to revise accordingly the path to your local clone of Just_universal **
source("~/liuyanguu/Just_universal/code/xgboost_cvtune.R")
source("~/liuyanguu/Just_universal/code/xgboost_cv_RFE.R")

# Parameters ####
# the aer site info. file:
aer_stn_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/aeronet_locations_v3.txt"
# aer_data_path = '~/qnap_geo/aeronet/All_Sites_Times_All_Points_AOD20_20180603.dat'
# aer_data_path = '/scratch/temp/All_Sites_Times_All_Points_AOD20_20180603.dat' 
# the aer data files:
aer_files_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
date_start = "2018-01-01" # do not include in sel_data_bytime if don't want to limit
date_end   = "2018-12-31" # do not include in sel_data_bytime if don't want to limit
sats = c("terra", "aqua")

# MCD19 data load and save output: 
mcd19path = "/data-coco/mcd19/fst/nemia/2018/" # now there is only 2018 data 
mcd19path_CONUS = "/data-coco/mcd19/fst/conus/2018/"
MCD19_T_files <- list.files(path = mcd19path_CONUS, pattern = "T", full.names = T)
MCD19_A_files <- list.files(path = mcd19path_CONUS, pattern = "A", full.names = T)
out_dir = "/data-coco/mcd19/fst/conus/2018_AOD_newVar/"
# how many MCD days (files) to work on:
MCD_files_i = c(1:length(MCD19_T_files))

# for reporting purpose
y_var = "diff_AOD"
y_var_pred = "diff_AOD_pred"
# how many features to print in scatterplots
n_vars = 10

# point to cache: 
ssd_cache = new_cache(path = "/scratch/cache/aeronet_drake") 
```

# Plan part 1

```{r}
# CONUS Drake Plans  ------------------------------------------------------------
source(here("R/drake_funcs.R"))
conus_plan <- drake_plan(
  # 1.  Aeroent ---------------------------------------------------------------------

@@

@@

  # 3. WPred Join MCD19 -----------------------------------------------------------
  # mcd is the large MCD19 dataset, limit to `refsub`:nearby(300km) when readin to 
  # control total lines, unable to read in all data together, too many rows
  mcd = target(read_mcd19_one(i, sat = sat, file_in(mcd19path_CONUS)), 
               transform = cross( i = !!MCD_files_i, 
                                  sat = !!sats),
               format = "fst"),
  # join is the rolling-joined dataset with aeronet
  join = target(rolling_join(mcd_sat = mcd, aer_data_wPred = wPred),
                transform = cross(mcd, wPred, .id = mcd)),

  # For calculating new variables
  # selected aeronet 
  sel_aer = target(get_conus_aer_used(conus_aer, sel_aer_region),
                  transform = map(sel_aer_region, .id = FALSE)),

  # buffers 
  ref = target(ref_in_buffer(sel_aer, refgrid),
             transform = map(sel_aer)),
  # sites
  p = target(get_site_list(sel_aer), transform = map(sel_aer)),
  n = target(get_site_name(sel_aer), transform = map(sel_aer)),

  # mapping 
  # this way doesn't work:
  # sub_grid = target(select_refgrid_subset(ref_list0 = ref, point0 = p, name0 = n),
  #              transform = map(ref, p, n))
  sub_grid = target(purrr_pmap(ref = ref, p = p, n = n),
               transform = map(ref, p, n, .id = FALSE),
               format = "fst"),
  # calculate new vars to obtain the dataset ready for modeling for terra and aqua, using join, mcd, sub_grid 
  new_var = target(aod_MODIS_newVars(aod_join_MODIS = join, 
                                 MODIS_all = mcd, refDT_sub = sub_grid),
               transform = map(join, mcd, .id = c(join)),
               format = "fst"),
  # deliver two rowbinded dataset: for terra and aqua 
  # I am not sure what this way doesn't work: 
  # binded = target(rbindlist(new_var),
  #                 transform = combine(new_var, .by = c(i, sat))),
  # `.id_chr` let you save the file using target names:
  out = target(write_new_var(new_var, id = .id_chr), transform = map(new_var))
)

conus_plan
nrow(conus_plan)
#drake_plan_source(conus_plan)
if(nrow(conus_plan)<100){
nemia_config <- drake_config(conus_plan, cache = ssd_cache) # show the dependency
# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
vis_drake_graph(nemia_config, ncol_legend = 0)
}
```

# Plan part 2

```{r}
# drake plan to run cv and output results 
conus_plan_cv <- drake_plan(
  # 5. CV -----------------------------------------------------------
  new_var = target(read_new_var(sat), transform = map(sat = !!sats)),
  cv_results = target(run_cv(dt = new_var), transform = map(new_var, .id = sat)),

@@

  shap_long = target(get_SHAP_long(cv_results),
                  transform = map(cv_results,.id = FALSE)),
  plotSHAP1 = target(plot_SHAP(shap_long),
                 transform = map(shap_long,.id = FALSE)),
  plotSHAP2 = target(plot_SHAP_scatter(cv_results, shap_long),
                 transform = map(cv_results, shap_long,.id = FALSE)),
  report = rmarkdown::render(
    knitr_in("AOD_CONUS_Report.Rmd"),
    output_file = file_out("AOD_CONUS_Report.html"))
)
nemia_config <- drake_config(conus_plan_cv, cache = ssd_cache) # show the dependency
# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
vis_drake_graph(nemia_config)
```

# Run the two plans

```{r, eval = F}
# make(conus_plan, cache = ssd_cache)

time0 <- Sys.time()
future::plan(future::multiprocess)
options(future.globals.maxSize= 100*1024^3) 
make(conus_plan, cache = ssd_cache,
     parallelism = "future", jobs = 20)
Sys.time() - time0

make(conus_plan_cv, cache = ssd_cache)
